Component response rate variation underlies the stability of complex systems

### 2020-03-11 notebook/ms.Rmd
Add the mention of C values into the results for rho versus eigenvalues

Ten thousand numerical simulations of $\mathbf{M}$ revealed that $\sigma^{2}_{\gamma} > 0$ results in a nonlinear relationship between $\rho$ and $\Re(\lambda_{max})$, which can sometimes increase the stability of the system. Figure 2 shows a comparison of $\Re(\lambda_{max})$ across $\rho$ values when $\sigma^{2}_{\gamma} = 0$ versus $\sigma^{2}_{\gamma} = 1/3$ given $S = 25$ and $\sigma = 0.4$. For $-0.4 \leq \rho \leq 0.7$ (shaded region of Figure 2), expected $\Re(\lambda_{max})$ was lower given $\sigma^{2}_{\gamma} = 1/3$. For $\rho \geq -0.1$, the range of $\Re(\lambda_{max})$ values observed given $\sigma^{2}_{\gamma} = 1/3$ also had a lower bound, which included negative values for $\rho = -0.1$ and $\rho = 0$. Hence, across a wide range of system correlations, variation in the response rate of system components had a stabilising effect.

The stabilising effect of $\sigma^{2}_{\gamma}$ across $\rho$ increased with increasing $S$. Figure 3 shows 10000 numerical simulations of $\mathbf{M}$ for $\sigma^{2}_{\gamma} = 0$ versus $\sigma^{2}_{\gamma} = 1/3$ across increasing $S$ given $\sigma = 0.2$ ($\sigma$ has been lowered here to better illustrate the effect of $S$; note that now given $S = 25$, $1 = \sigma\sqrt{SC}$). For relatively small systems ($S \leq 25$), positive $\sigma^{2}_{\gamma}$ never decreased the expected $\Re(\lambda_{max})$. But as $S$ increased, the curvilinear relationship between $\rho$ and $\Re(\lambda_{max})$ resulted in $\mathbf{M}$ with lower expected $\Re(\lambda_{max})$ for low magnitudes of $\rho$. In turn, as $S$ increased, and systems became more complex, positive $\sigma^{2}_{\gamma}$ increased the proportion of numerical simulations that were observed to be stable (see below).

>>>>

Ten thousand numerical simulations of $\mathbf{M}$ revealed that $\sigma^{2}_{\gamma} > 0$ results in a nonlinear relationship between $\rho$ and $\Re(\lambda_{max})$, which can sometimes increase the stability of the system. Figure 2 shows a comparison of $\Re(\lambda_{max})$ across $\rho$ values when $\sigma^{2}_{\gamma} = 0$ versus $\sigma^{2}_{\gamma} = 1/3$ given $S = 25$, $C = 1$, and $\sigma = 0.4$. For $-0.4 \leq \rho \leq 0.7$ (shaded region of Figure 2), expected $\Re(\lambda_{max})$ was lower given $\sigma^{2}_{\gamma} = 1/3$. For $\rho \geq -0.1$, the range of $\Re(\lambda_{max})$ values observed given $\sigma^{2}_{\gamma} = 1/3$ also had a lower bound, which included negative values for $\rho = -0.1$ and $\rho = 0$. Hence, across a wide range of system correlations, variation in the response rate of system components had a stabilising effect.

The stabilising effect of $\sigma^{2}_{\gamma}$ across $\rho$ increased with increasing $S$. Figure 3 shows 10000 numerical simulations of $\mathbf{M}$ for $\sigma^{2}_{\gamma} = 0$ versus $\sigma^{2}_{\gamma} = 1/3$ across increasing $S$ given $C = 1$ and $\sigma = 0.2$ ($\sigma$ has been lowered here to better illustrate the effect of $S$; note that now given $S = 25$, $1 = \sigma\sqrt{SC}$). For relatively small systems ($S \leq 25$), positive $\sigma^{2}_{\gamma}$ never decreased the expected $\Re(\lambda_{max})$. But as $S$ increased, the curvilinear relationship between $\rho$ and $\Re(\lambda_{max})$ resulted in $\mathbf{M}$ with lower expected $\Re(\lambda_{max})$ for low magnitudes of $\rho$. In turn, as $S$ increased, and systems became more complex, positive $\sigma^{2}_{\gamma}$ increased the proportion of numerical simulations that were observed to be stable (see below).
